WORKING TITLE: The adaptive value of green nest material across birds: A systematic review and meta-analysis

### 2025-07-05 manuscript_GNM.qmd
Some info added to manuscript.. trying to fix supplementary material.
Updated Introduction/Methods completed Added all the codes and data files to the folder. Can publish directly from here for later..

@@

```{r Setup, include=FALSE, echo=FALSE}
devtools::install_github("daniel1noble/orchaRd", ref = "main", force = TRUE)
pacman::p_load(readr,tidyverse,metafor,orchaRd,patchwork)
source(here::here("functions/extractor_metareg.R"))
```

@@

# Systematic search and abstract screening
unique_reference_list <- read_csv("data/unique_reference_list.csv")
full_reference_before_deduplication<-read_csv("data/full_reference_before_deduplication.csv")
selected_abstract_screening<-read_csv("data/selected_abstract_screening.csv")

# For updated search
unique_reference_list_repeat <- read_csv("data/unique_reference_repeat_12082024.csv")
full_reference_before_deduplication_repeat<-read_csv("data/full_reference_repeat_12082024.csv")
selected_abstract_screening_repeat<-read_csv("data/selected_abstract_screening_repeat.csv")

# For fulltext screening
excluded_fulltext <- read_csv("data/excluded_fulltext.csv")
```

We pre-registered our study protocol on Open Science Framework [@dimri2024] prior to data exploration and analysis. The pre-registration details our objectives, a priori hypotheses and predictions, methods and planned data analyses. Unless stated otherwise, we adhered to the registered protocol. We followed the PRISMA-EcoEvo guidelines [@odea2021] for reporting this study (@figsupp-checklist). All resources associated with this study, including data and code, are available at \[<https://github.com/shreyadimri/Green_Nest_Material>\] (insert Zenodo citation later).

We conducted a systematic search on Web of Science Core Collection and Scopus on 07.09.2022. The keyword search query was designed as a three-block query to included terms related to the predictor/treatment (("green\*" OR "herb\*" OR "aromatic\*") AND "nest\*"), our population of interest ("bird\*" OR "aves" OR "avian" OR "ornithol\*" OR "passerine\*" OR "passeriform\*" OR "songbird\*" OR list of all bird genera) and to limit it to only experimental studies only ("experiment\*" OR "manipulat\*"). The complete search strings tailored to each specific platform are provided in @figsupp-string. We searched through titles, abstracts and keywords in English-language only with no restrictions on publication date or study type. We validated our search using an initial library of `{r} nrow(own_library)` relevant articles that we had previously identified (@figsupp-own-library), ensuring all were retrieved. Additionally, we used the R package 'litsearchr' v1.0.0 (@grames2019) to confirm our keyword selection, which did not yield any new relevant terms. We found `{r} nrow(unique_reference_list)` unique articles after removing `{r} nrow(full_reference_before_deduplication)-nrow(unique_reference_list)` duplicates using R package 'revtools' v0.4.1 (@westgate2019).

@@

All titles and abstracts were equally allocated and screened independently by atleast two screeners (SD, TR, JMGS) using the R package `revtools` following a pre-designed decision tree (@figsupp-decision-tree). We included all studies on birds that experimentally manipulated the green nest material (both aromatics or herbs). `{r} nrow(selected_abstract_screening)` articles passed the title-and-abstract screening and were further evaluated during the full-text screening. Each article was screened by a single screener (SD, TR, JMGS), with another screener verifying the procedure. We used the same criteria as above (see @figsupp-decision-tree) with an additional evaluation of whether the variable of interest in the article was in our list of fitness proxies of interest (see @figsupp-fitness-proxies). `{r} nrow(excluded_fulltext)` articles were excluded. All conflicts in screening decisions (\~1% disagreement for title-and-abstract screening, and none for full-text screening) were resolved through discussion.

To update our systematic lietrature search, we repeated our search on 12.08.2024. This updated search found `{r} nrow(unique_reference_list_repeat)` unique articles after removing ``` {r} nrow(full_reference_before_deduplication_repeat)-``nrow(unique_reference_list_repeat) ``` duplicates. Only one observer (SD) performed the title-and-abstract and full-text screening using the same decision trees as before and found `{r} sum(selected_abstract_screening_repeat$screened_abstracts=="selected", na.rm=T)` additional eligible article. Lastly, we identified `{r} nrow(selected_abstract_screening_repeat)-(sum(selected_abstract_screening_repeat$screened_abstracts=="selected", na.rm=T))` additional eligible articles. In all, X <!--# Fill the no. of articles that are included --> articles were eligible for our meta-analysis (more details in @figsupp-prisma-flowchart).

@@

dataset_after_cleaning <- merged_data_extraction %>% filter(fulltext_screening!="excluded")
included_papers_no<-length(unique(dataset_after_cleaning$paper_ID))

@@

excluded_proxies<- dataset_after_cleaning %>% filter(proxy_decision=="exclude")
SAonly_proxies<- dataset_after_cleaning %>% filter(proxy_decision=="senstivity-analysis" | proxy_decision=="sensitivity-analysis")
```

The data of each article was extracted evenly by one of the three authors (SD, TR, JMGS), with another screener verifying the procedure. We extracted the primary data (i.e., measures of central tendency, variance and sample size) from text whenever possible. In cases where primary data was only available in figures (e.g., boxplots), we used `metaDigitise` v.1.0.1 (@pick2018). If primary data was not available, we extracted any inferential statistics (t value, F value) that could be transformed into an effect size. In cases where the values could not be extracted from the article but datasets were provided, we calculated the means, SD and sample sizes from the provided raw data. When data was presented in different sources (e.g. in a boxplot and as an inferential statistics), we prioritized the source having the maximum resolution (the boxplot, in this example).

We extracted the means, variances and sample size for each control-treatment pairwise comparison. Controls included non-aromatic materials such as grass or moss, lesser amounts of aromatic material, or no material at all. Treatments consisted of aromatic materials added by authors, birds, or both. In cases comparing non-aromatic material to a blank nest, the non-aromatic material was considered the treatment. We recorded the type of comparison being made as a moderator.

For each study we also extracted information related to the study subjects (species of birds, location of the study population), the methodology of the study (plant species used as treatment and control, fitness proxy measured) and information about the extracted data (location or source of the data recorded). We categorized the fitness proxy measured into 6 broad categories - physiology, morphology, reproduction, behaviour, parasite and pathogenic load and phenology. For parasites or pathogens, we categorised them into arthropods or micro-organisms. For a complete list of variables extracted see @figsupp-variables . We did not include some traits in all our analyses becaause it was unclear how they related to fitness (k=`{r} nrow(excluded_proxies)` , n=`{r} length(unique(excluded_proxies$paper_ID))`;see @figsupp-excluded-proxies for the list and reason of exclusion). We also had to exclude some effect sizes (n=?, k=?) because of the low sample size per group (less than x).

To obtain missing data (e.g., non-reported results, missing sample sizes, etc.) and to locate grey literature (i.e. unpublished data), we wrote to the corresponding author of the articles that passed full-text screening using a standardized email template (\@figsupp-email). We contacted four authors for missing data, while all corresponding authors were asked for any potential unpublished data that could be included in our meta-analysis. All four authors responded to our email but only two could provide the data as requested; one did not have the data any more and one did not give a conclusive response. In addition, one author provided data for an unpublished study (see description in @figsupp-unpublished-study) and one of the authors of this study (MO) provided data for an unpublished study. In all, we extracted data for `{r} included_papers_no - 2` published articles and 2 unpublished studies.

@@

inferential_stats<-dataset_after_cleaning%>%filter(statistics_type=="t-test"|statistics_type=="ANCOVA")

contingency_table<-dataset_after_cleaning%>%filter(proxy_decision=="contingency table")

no_ES<-dataset_after_cleaning%>%filter(proxy_comment == "use 0 as ES")

negative_lnRR<-dataset_after_cleaning%>%
  filter(measure_central_tendency_experiment<=0 | measure_central_tendency_control<=0)

We calculated two effect sizes for each pairwise comparison: a log response ratio (lnRR) and a standardised mean difference with heteroskedasticity correction (SMDH), using the `escalc` function (`measure= ROM` and `SMDH` respectively, `vtype= "LS"`) in the R package `metafor` (@metafor). We specified the treatment group as the numerator and the control group as the denominator for lnRR and treatment minus control for SMDH, such that positive values indicate the 'fitness' value increased in the treatment group. In cases such as measures of parasite load or prevalence, laying or hatching dates, courtship time, scab score etc., the effect size sign was inverted to ensure comparability across all fitness proxies. In all cases, the effect sizes were coded such that the fitness proxies were positively correlated with the overall fitness.

For inferential statistics (e.g., *t* value, *F* value), we estimated SMD values (k=`{r} nrow(inferential_stats)`, n=`{r} length(unique(inferential_stats$paper_ID))`) using the equations provided by @thehand2019. For dichotomous variables (e.g., number of successful nests, k= `{r} nrow(contingency_table)`, n= `{r} length(unique(contingency_table$paper_ID))`), we calculated the log odds ratio (lnOR) using the `escalc` function (`measure=OR`) from `metafor` and subsequently converted it to SMD using equations from @thehand2019. Since lnRR cannot be calculated from inferential statistics, or dichotomous variables, or for comparisons including zero or negative values (k=`{r} nrow(negative_lnRR)`, n=`{r} length(unique(negative_lnRR$paper_ID))`), the number of effect sizes included in the lnRR (k=?, n=?) subset is smaller than that of SMDH (k=?,n=?).

When calculating each effect size's associated sampling variance, we prioritised the number of nests in each group over the number of individuals as the sample size, whenever possible to reduce the effect of genetic relatedness non-independence. In cases where the control or treatment group was used repeatedly across multiple comparisons, sample sizes were adjusted by dividing them by the number of times each group was used in a comparison to account for shared group non-independence. For the cases where the authors of an original article reported no statistically significant difference between two groups but did not provide numerical values (k=`{r} nrow(no_ES)`, n=`{r} length(unique(no_ES$paper_ID))`), we assigned zero as the effect size for that comparision and imputed its SD using the missing-case approach described in @nakagawa2022.

@@

We performed all analyses in R version 4.3.0 (@positteam2023). All multilevel meta-analytical models and meta-regressions were fit using `metafor v.4.8-0` (@metafor). Sampling variances were modelled as variance-covariance matrices assuming a 0.5 correlation (*ρ*) between sampling variances from the same study [@noble2017] using the `vcalc` function from `metafor`. (See Supplementary Appendix @figsupp-correlation for sensitivity analyses using different values of *ρ*).

We considered the following levels of non-independence for our analyses: (a) multiple effect sizes collected from same study (Paper ID), (b) bird species (Species ID) and (c) geographical locations (population ID), (d) some sub-populations (Experiment ID; e.g. birds studied at the same location for multiple years or birds studied at the same location separated by distance) and (e) group of individuals (Group ID; eg. males and females in a population, first clutch and second clutch) were represented in our data several times, and (f) some traits were collected on the same group of animals repeatedly (Repeated trait ID; e.g. chick mass on day 7, 14 and 21).

@@

To estimate the effect of green nest material addition on fitness across birds, we used two effect sizes: lnRR and SMDH. Thus, throughout, we fitted two separate sets of model for each effect size. First, we fitted multilevel intercept-only meta-analytical model to test the overall effect, its uncertainty and heterogeneity. We used the `rma.mv` function from the `metafor` package using the 'REML' method and a *t* distribution for calculating test statistics and confidence intervals for the fixed effects. We also conducted a sensitivity analysis including some traits for which the relationship with fitness was unclear, to test the robustness of our results (@figsupp-MLMA-excluded; k=`{r} nrow(SAonly_proxies)` , n=`{r} length(unique(SAonly_proxies$paper_ID))`; see @figsupp-SAonly_proxies). For all intercept-only models, we estimated the total amount of absolute heterogeneity ($𝜎^2$; the variation not explained by the sampling error), its sources (or relative heterogeneity; $I^2$), mean-standardised metric ($CVH2$) and variance-mean-standardised metric of heterogeneity ($M2$) following the pluralistic approach suggested by @yang2025.

<!--# You have to correct the numbers for contingency table. It does not include the one data point from unpublished results -->

@@

To test the relative importance of the two main hypotheses suggested for green nest material: Courtship hypothesis and Parental care hypothesis (combining nest protection and drug hypothesis), we ran a multilevel meta-regression with 'hypothesis' (Courtship hypothesis, Parental care hypothesis, Both) as a moderator. For this and all meta-regressions, we estimated $R_{marginal}^2$, which corresponds to the percentage of heterogeneity explained by the moderators (@nakagawa2012).

@@

We performed additional pre-registered multi-level meta-regressions for which we did not have clear or strong directional predictions on specific subsets of data. To understand the effect of green nest material on the parasite load depending on the type of parasites, we used effect sizes on parasites and pathogens (n=?, k=?). We ran a meta-regression with a type of parasite as a moderator (levels: arthropod, micro-organism). We also ran a meta-regression with with the entire dataset set using bird species as a moderator (levels: Cyanistes caeruleus, Sturnus unicolor, Tachycineta bicolor, Sturnus vulgaris, Parus major, Passer cinnamomeus, Buteo buteo) to understand if the effect of green nest material on fitness depended on the system being studied. To test the effect of time of addition of green nest material on the fitness outcome (levels: before egg hatching, after egg hatching, continuously throughout the nesting phase) we ran a meta-regression with Time of addition of green nest material as the moderator. We explored the effect of green nest material on fitness depending on the type of trait studies (levels:physiology, morphology, reproduction, behaviour, parasite and pathogenic load, phenology) using a meta-regression with the trait type as a moderator. We explored the impact of methodological differences on the relationship between green nest material and fitness. To understand the effect of type of experimental design (i.e. whether the authors compared - non-aromatic vs. aromatic, no added material vs. aromatic, no added material vs. non-aromatic), we ran a meta-regression with the type of experimental design as a moderator.

@@

```{r review-results}
dataset_analysis <- read_csv("data/03_data_cleaning/dataset_analysis.csv")

@@

We obtained `{r} nrow(dataset_analysis)` effect sizes from `{r} length(unique(dataset_analysis$paper_ID))` studies conducted on `{r} length(unique(dataset_analysis$bird_species))` species across `{r} length(unique(dataset_analysis$population_ID))` geographical locations. About `{r} round(sum(birds_info$perc[birds_info$perc>15]))`% of the effect sizes came from three bird species (`{r} birds_info$bird_species[birds_info$perc > 15][1]`, k = `{r} birds_info$n[birds_info$perc > 15][1]`; `{r} birds_info$bird_species[birds_info$perc > 15][2]`, k = `{r} birds_info$n[birds_info$perc > 15][2]`; `{r} birds_info$bird_species[birds_info$perc > 15][3]`, k = `{r} birds_info$n[birds_info$perc > 15][3]`). Our dataset included two unpublished studies (k=XX effect sizes), with the remaining studies being published between `{r} min((dataset_analysis$year_publication),na.rm = T)` and `{r} max((dataset_analysis$year_publication),na.rm = T)`. The mean (±SD) number of effect sizes per study was `{r} as.numeric(format(mean(ES_per_study$n), digit=3))` ± `{r} as.numeric(format(sd(ES_per_study$n), digit=2))`(range: `{r} min(ES_per_study$n)`--`{r} max(ES_per_study$n)`) and the median and mean sample size per effect size were `{r} median(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment))` and `{r} as.numeric(format(mean(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment)), digit=3))` respectively (range: `{r} min(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment))`--`{r} max(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment))`; `{r} nrow(dataset_analysis%>%select(paper_ID,effective_n_control,effective_n_experiment, lnRR, SMDH)%>%filter(effective_n_experiment<2|effective_n_control<2))` effect sizes had a sample size of less than 2 in each group).

@@

Our systematic review and meta-analysis found a small positive increase in fitness benefits in the bird nests with increased green nest material, though the strength of these effects depended on the effect size metric used. SMDH showed a larger effect with the There is high heterogeneity among the effect sizes, emphasizing the context-dependence of this benefit. Importantly, our results reveal that the commonly studied mechanistic hypotheses of green nest material---whether through sexual signaling, parasite reduction, or direct nestling benefits---lack consistent empirical support across studies.

@@

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\--

## Supplementary Appendix

@@

<details>

<summary>Supplementary materials</summary>

{{< include supplementary_appendix.qmd >}}

</details>

>>>>

# Abstract

According to PRISMA Checklist.. I must:

-   Summarise the aims and scope of the review

-   Describe the data set

-   State the results of the primary outcome

-   State conclusions

-   State limitations

@@

```{r Setup, include=FALSE, echo=FALSE}
devtools::install_github("daniel1noble/orchaRd", ref = "main", force = TRUE)
pacman::p_load(readr,tidyverse,metafor,orchaRd,patchwork,english)
source(here::here("functions/extractor_metareg.R"))
```

We pre-registered our study protocol on Open Science Framework [@dimri2024] prior to data exploration and analysis. The pre-registration details our objectives, a priori hypotheses and predictions, methods and planned data analyses. Unless stated otherwise, we adhered to the registered protocol. We followed the PRISMA-EcoEvo guidelines [@odea2021] for reporting this study (@figsupp-checklist and @figsupp-prisma-flow). All resources associated with this study, including data and code, are available at \[<https://github.com/shreyadimri/Green_Nest_Material>\] (insert Zenodo citation later).

@@

# Systematic search and abstract screening
unique_reference_list <- read_csv("data/01_systematic_search/02_reference_data/unique_reference_list.csv")
full_reference_before_deduplication<-read_csv("data/01_systematic_search/02_reference_data/full_reference_before_deduplication.csv")
selected_abstract_screening<-read_csv("data/01_systematic_search/03_title_and_abstract_screening/selected_abstract_screening.csv")

# For updated search
unique_reference_list_repeat <- read_csv("data/01_systematic_search/04_search_repeat_12082024/unique_reference_repeat_12082024.csv")
full_reference_before_deduplication_repeat<-read_csv("data/01_systematic_search/04_search_repeat_12082024/full_reference_repeat_12082024.csv")
selected_abstract_screening_repeat<-read_csv("data/01_systematic_search/04_search_repeat_12082024/selected_abstract_screening_repeat.csv")

# For fulltext screening
excluded_fulltext <- read_csv("data/01_systematic_search/03_title_and_abstract_screening/excluded_fulltext.csv")
```

We conducted a systematic search on Web of Science Core Collection and Scopus on 07.09.2022 (see update below). The keyword search query was designed as a three-block query to include terms related to our treatment of interest (("green\*" OR "herb\*" OR "aromatic\*") AND "nest\*"), population of interest ("bird\*" OR "aves" OR "avian" OR "ornithol\*" OR "passerine\*" OR "passeriform\*" OR "songbird\*" OR list of all bird genera) and to limit the search results to experimental studies only ("experiment\*" OR "manipulat\*"). The complete search strings tailored to each specific platform are provided in @figsupp-string. We searched through titles, abstracts and keywords in English-language only with no restrictions on publication date or study type. We validated our search using an initial library of `{r} nrow(own_library)` relevant articles that we had previously identified (@figsupp-own-library), ensuring all were retrieved. Additionally, we used the R package 'litsearchr' v1.0.0 [@grames2019], which did not yield any new relevant terms, to confirm our keyword selection. After removing `{r} nrow(full_reference_before_deduplication)-nrow(unique_reference_list)` duplicates using R package 'revtools' v0.4.1 [@westgate2019], we found `{r} nrow(unique_reference_list)` unique records.

@@

All titles and abstracts were equally allocated and screened independently by atleast two screeners (SD, TR, JMGS) using the R package `revtools` following a pre-designed decision tree with all inclusion/exclusion criteria (@figsupp-decision-tree). We included all studies on birds that experimentally manipulated the green nest material (both aromatics or herbs). `{r} tools::toTitleCase(as.character(english::english(nrow(selected_abstract_screening))))` articles passed the title-and-abstract screening and were further evaluated during the full-text screening, where each article was independently screened by a single screener (SD, TR, JMGS), with another screener verifying the procedure. We used the same criteria as above (@figsupp-decision-tree) with an additional evaluation of whether the variable of interest in the article was in our list of fitness proxies of interest (see @figsupp-fitness-proxies). `{r} paste0(toupper(substring(x <- as.character(english::english(nrow(selected_abstract_screening) - nrow(excluded_fulltext))), 1, 1)), substring(x, 2))` passed the full-text screening and were, thus, eligible for our meta-analysis. All conflicts in screening decisions (\~1% disagreement for title-and-abstract screening, and none for full-text screening) were resolved through discussion.

We updated our systematic literature search by repeating our search on 12.08.2024. We found `{r} nrow(unique_reference_list_repeat)` unique articles after removing `{r} nrow(full_reference_before_deduplication_repeat)-nrow(unique_reference_list_repeat)` duplicates. Only one observer (SD) performed the title-and-abstract and full-text screening using the same decision trees as before (@figsupp-decision-tree), and found `{r} sum(selected_abstract_screening_repeat$screened_abstracts=="selected", na.rm=T)` additional eligible article for our meta-analysis. Lastly, we identified `{r} nrow(selected_abstract_screening_repeat)-(sum(selected_abstract_screening_repeat$screened_abstracts=="selected", na.rm=T))` additional eligible articles through other sources (i.e., identified through a poster by a collaborator and through an independent recommendation by a colleague). In all, `{r} nrow(selected_abstract_screening)-nrow(excluded_fulltext)+nrow(selected_abstract_screening_repeat)` articles were eligible for our meta-analysis (more details in @figsupp-prisma-flowchart).

@@

dataset_after_cleaning <- read_csv("data/03_data_cleaning/dataset_after_cleaning.csv", show_col_types = F)
included_papers_no<-length(unique(dataset_after_cleaning$paper_ID))

@@

excluded_proxies<- merged_data_extraction %>%
  filter(fulltext_screening == "included" |
         is.na(fulltext_screening) |
         fulltext_screening == "NA") %>%
  filter(proxy_decision== "exclude")

dataset_analysis <- read_csv("data/04_data_analysis/dataset_analysis.csv", show_col_types = F)

SAonly_proxies<- dataset_analysis %>% filter(proxy_decision=="sensitivity-analysis")

inferential_only<- dataset_analysis%>%filter(statistics_type=="t-test"|statistics_type=="ANCOVA")

contingency_table<-dataset_analysis%>%filter(proxy_decision=="contingency table")

raw_data<-c("GNM_001_rep_table3","GNM_010","GNM_069","GNM_237") 
#GNM_018 also provides open data but we ended up not needing it

published_studies <- dataset_analysis %>%
  filter(!str_detect(paper_ID, "Unpublished"))

unpublished_studies <- dataset_analysis %>%
  filter(str_detect(paper_ID, "Unpublished"))

The data of each article was extracted evenly by one of the three authors (SD, TR, JMGS), with another screener verifying the procedure. We prioritised extracting the primary data (i.e., measures of central tendency, variance and sample size) from text and tables, or, if that was not possible, from figures (e.g., boxplots) using the R package `metaDigitise` v.1.0.1 (@pick2018). If primary data were not available, we extracted inferential statistics such as *t* and *F* values (k = `{r} nrow(inferential_only)`, n = `{r} length(unique(inferential_only$paper_ID))`) or descriptive statistics such as contingency tables (k = `{r} nrow(contingency_table)`, n = `{r} length(unique(contingency_table$paper_ID))`), that could be transformed into an effect size. Last, if the previous steps failed but the article has openly shared datasets (n = `{r} length(raw_data)`), we calculated the means, SD and sample sizes from them. **When data was presented in different sources (e.g. in a boxplot and as an inferential statistics), we prioritized the source having the maximum resolution (the boxplot, in this example).**

We extracted the means, variances and sample size for each control-treatment pairwise comparison. Controls included non-aromatic materials such as grass or moss, lesser amounts of aromatic material, or no material at all. Treatments consisted of aromatic materials added by authors, birds, or both. For cases that compared non-aromatic material to where authors did not add any additional nest material and any material added by the birds was removed, the non-aromatic material was considered the treatment. We recorded the type of comparison being made as a moderator. For each study, we also extracted information related to the study subjects (species of birds, location of the study population), the methodology of the study (plant species used as treatment and control, fitness proxy measured) and the source from which we extracted data (e.g., figures, tables, raw data files, or supplementary materials). We categorised the fitness proxies measured into 6 broad categories - physiology, morphology, reproduction, behaviour, parasite and pathogenic load, and phenology. For studies testing whether green nest material affects the number of parasites or pathogens found in offspring and/or nests, we categorised them into arthropods or microorganisms. For a complete list of variables extracted see @figsupp-variables . We did not include some traits in all our analyses because it was unclear how they related to fitness (k = `{r} nrow(excluded_proxies)` , n = `{r} length(unique(excluded_proxies$paper_ID))`; see @figsupp-excluded-proxies for the list and reasons of exclusion).

To obtain missing data (e.g., non-reported results, missing sample sizes, etc.), we contacted four authors. All four authors responded to our email, but only two provided the data as requested; one did not have the data any more, and one did not give a conclusive response. Additionally, to locate grey literature (i.e., unpublished data), we wrote to the corresponding author(s) of all the articles that passed full-text screening (n = `{r} (nrow(selected_abstract_screening) - nrow(excluded_fulltext))`) to ask for any potential unpublished data that could be included in our meta-analysis. One corresponding author provided data for an unpublished study (see description in @figsupp-unpublished-study) and one of the authors of this study (MO), provided data for another unpublished study. In all, we extracted data for `{r} included_papers_no - 2` published articles and 2 unpublished studies. Author correspondance was performed using a standardised email template (@figsupp-email). In all, we extracted data for `{r} length(unique(published_studies$paper_ID))` published articles (k = `{r} nrow(published_studies)`) and `{r} length(unique(unpublished_studies$paper_ID))` unpublished studies (k = `{r} nrow(unpublished_studies)`).

@@

no_ES<-dataset_analysis%>%filter(proxy_comment == "use 0 as ES")

negative_lnRR<-dataset_analysis%>%
  filter(measure_central_tendency_experiment<=0 | measure_central_tendency_control<=0)

low_sample<-dataset_analysis%>%filter(is.na(SMDH_sign))

dataset_lnRR <- read_csv("data/04_data_analysis/dataset_lnRR.csv", show_col_types = F)

dataset_SMDH <- read_csv("data/04_data_analysis/dataset_SMDH.csv", show_col_types = F)
```

We calculated two effect sizes for each pairwise comparison: a log response ratio (lnRR) and a standardised mean difference with heteroskedasticity correction (SMDH), using the `escalc` function (`measure= ROM` and `SMDH` respectively, `vtype= "LS"`) in the R package `metafor` [@metafor]. We specified the treatment group as the numerator and the control group as the denominator for lnRR and treatment minus control for SMDH, such that positive values indicate the 'fitness' value increased in the treatment group. In cases such as measures of parasite load or prevalence, laying or hatching dates, courtship time, scab score etc., the effect size sign was inverted to ensure comparability across all fitness proxies. In all cases, the effect sizes were coded such that the fitness proxies were positively correlated with the overall fitness. SMDH could only be calculated when there were enough data points in both treatment and control groups to estimate the pooled variance— specifically, when the degrees of freedom (i.e., n1+n2−2) was greater than 1 (excluded k = `{r} nrow(low_sample)`, n = `{r} length(unique(low_sample$paper_ID))`)

For inferential statistics (e.g., *t* value, *F* value), we estimated SMD values (k = `{r} nrow(inferential_only)`, n = `{r} length(unique(inferential_only$paper_ID))`) using the equations provided by @thehand2019. For dichotomous variables (e.g., number of successful nests, k = `{r} nrow(contingency_table)`, n = `{r} length(unique(contingency_table$paper_ID))`), we calculated the log odds ratio (lnOR) using the `escalc` function (`measure=OR`) from `metafor` and subsequently converted it to SMD using equations from @thehand2019. Since lnRR cannot be calculated from inferential statistics, or dichotomous variables, or for comparisons including zero or negative values (k = `{r} nrow(negative_lnRR)`, n = `{r} length(unique(negative_lnRR$paper_ID))`), the number of effect sizes included in the lnRR (k = `{r} nrow(dataset_lnRR)`, n = `{r} length(unique(dataset_lnRR$paper_ID))`) subset is smaller than that of SMDH (k = `{r} nrow(dataset_SMDH)`, n = `{r} length(unique(dataset_SMDH$paper_ID))`).

When calculating each effect size's associated sampling variance, we prioritised the number of nests in each group over the number of individuals as the sample size, whenever possible to reduce the effect of genetic relatedness non-independence. In cases where the control or treatment group was used repeatedly across multiple comparisons, sample sizes were adjusted by dividing them by the number of times each group was used in a comparison to account for shared group non-independence. For the cases where the authors of an original article reported no statistically significant difference between two groups but did not provide numerical values (k = `{r} nrow(no_ES)`, n = `{r} length(unique(no_ES$paper_ID))`), we assigned zero as the effect size for that comparison and imputed its sampling variance using the missing-case approach described in @nakagawa2022 for lnRR and using $\frac{n_1+n_2}{n_1×n_2}$ for SMD [@lajeunesse2013].

@@

We performed all analyses in R version 4.5.1 [@R-version]. All multilevel meta-analytical models and meta-regressions were fit using metafor v.4.8-0 [@metafor]. Sampling variances were modelled as variance-covariance matrices assuming a 0.5 correlation (*ρ*) between sampling variances from the same study [@noble2017] using the vcalc function from metafor. (See Supplementary Appendix @figsupp-correlation for sensitivity analyses using different values of *ρ*).

We considered the following levels of non-independence for our analyses: multiple effect sizes collected from (a) same study (Paper ID), (b) bird species (Species ID) and (c) geographical locations (population ID), (d) sub-populations within the same location (Experiment ID; e.g. birds studied at the same location for multiple years or birds studied at the same location separated by distance), (e) groups of individuals within a population (Group ID; eg. males and females in a population, first clutch and second clutch) were represented in our data several times, and (f) repeated measurement of the same trait within the same individuals (Repeated trait ID; e.g. chick mass on day 7, 14 and 21).

@@

To estimate the effect of green nest material addition on fitness across birds, we used two effect sizes: lnRR and SMD(H). Thus, throughout, we fitted two separate sets of model for each effect size. First, we fitted multilevel intercept-only meta-analytical model to test the overall effect, its uncertainty and heterogeneity. We used the `rma.mv` function from the `metafor` package using the 'REML' method and a *t* distribution for calculating test statistics and confidence intervals for the fixed effects. We also conducted a sensitivity analysis including some traits for which the relationship with fitness was unclear, to test the robustness of our results (@figsupp-MLMA-excluded; k = `{r} nrow(SAonly_proxies)` , n = `{r} length(unique(SAonly_proxies$paper_ID))`; see @figsupp-SAonly-proxies). For all intercept-only models, we estimated the total amount of absolute heterogeneity ($𝜎^2$; the variation not explained by the sampling error), its sources (or relative heterogeneity; $I^2$), mean-standardised metric ($CVH2$) and variance-mean-standardised metric of heterogeneity ($M2$) following the pluralistic approach suggested by @yang2025.

@@

To test the relative importance of the two main hypotheses suggested for green nest material: Courtship hypothesis and Parental care hypothesis (combining nest protection and drug hypothesis), we ran a multilevel meta-regression with 'hypothesis' (Courtship hypothesis, Parental care hypothesis, Both) as a moderator. For this and all meta-regressions, we estimated $R_{marginal}^2$, which corresponds to the percentage of heterogeneity explained by the moderators [@nakagawa2012].

@@

```{r}
parasite_pathogen<-dataset_after_cleaning%>%filter(trait_type=="Parasitic_and_pathogenic")
```

We performed additional pre-registered multi-level meta-regressions for which we did not have clear or strong directional predictions on specific subsets of data. To understand the effect of green nest material on the parasite load depending on the type of parasites, we used effect sizes on parasites and pathogens (k = `{r} nrow(parasite_pathogen)`, n = `{r} length(unique(parasite_pathogen$paper_ID))`). We ran a meta-regression with a type of parasite as a moderator (levels: arthropod, micro-organism). We also ran a meta-regression with the entire dataset set using bird species as a moderator (levels: *Buteo buteo*, *Cyanistes caeruleus*, *Passer cinnamomeus*, *Parus major*, *Sturnus unicolor*, *Sturnus vulgaris*, *Tachycineta bicolor*) to understand if the effect of green nest material on fitness depended on the system being studied. To test the effect of time of addition of green nest material on the fitness outcome (levels: before egg hatching, after egg hatching, continuously throughout the nesting phase), we ran a meta-regression with Time of addition of green nest material as the moderator. We explored the effect of green nest material on fitness depending on the type of trait studies (levels: physiology, morphology, reproduction, behaviour, parasite and pathogenic load, phenology) using a meta-regression with the trait type as a moderator. We explored the impact of methodological differences on the relationship between green nest material and fitness. To understand the effect of type of experimental design (i.e. whether the authors compared - non-aromatic vs. aromatic, no added material vs. aromatic, no added material vs. non-aromatic), we ran a meta-regression with the type of experimental design as a moderator.

@@

```{r review-results}

@@

We obtained `{r} nrow(dataset_analysis)` effect sizes from `{r} length(unique(dataset_analysis$paper_ID))` studies conducted on `{r} length(unique(dataset_analysis$bird_species))` species across `{r} length(unique(dataset_analysis$population_ID))` geographical locations. About `{r} round(sum(birds_info$perc[birds_info$perc>15]))`% of the effect sizes came from three bird species (`{r} birds_info$bird_species[birds_info$perc > 15][1]`, k = `{r} birds_info$n[birds_info$perc > 15][1]`; `{r} birds_info$bird_species[birds_info$perc > 15][2]`, k = `{r} birds_info$n[birds_info$perc > 15][2]`; `{r} birds_info$bird_species[birds_info$perc > 15][3]`, k = `{r} birds_info$n[birds_info$perc > 15][3]`). Our dataset included `{r} length(unique(unpublished_studies$paper_ID))` unpublished studies (k = `{r} nrow(unpublished_studies)`), with the remaining studies being published between `{r} min((dataset_analysis$year_publication),na.rm = T)` and `{r} max((dataset_analysis$year_publication),na.rm = T)`. The mean (±SD) number of effect sizes per study was `{r} sprintf("%.1f", mean(ES_per_study$n))` ± `{r} sprintf("%.1f",sd(ES_per_study$n))` (range: `{r} min(ES_per_study$n)`--`{r} max(ES_per_study$n)`) and the median and mean effective sample size (i.e. after accounting for shared-group non-independence) were `{r} sprintf("%.1f", median(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment)))` and `{r} sprintf("%.1f",mean(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment)))`, respectively (range: `{r} min(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment))`--`{r} max(c(dataset_analysis$effective_n_control,dataset_analysis$effective_n_experiment))`; `{r} nrow(dataset_analysis%>%select(paper_ID,effective_n_control,effective_n_experiment, lnRR, SMDH)%>%filter(effective_n_experiment<2|effective_n_control<2))` effect sizes had a sample size of less than 2 in each group).

@@

Summary of the results.. wrap up with a few points

Second paragraph

-   Discussing your results.. effect sizes

-   Difference between the hypothesis, whether you can do it or not

-   Experimental design difference.. importance of aromatic vs. Non-aromatic (are the same that aroma per say) and time of addition

    Third paragraph :

How much heterogeneity is explained..

Limitations and issues about the data.. mechanism, publication bias.. species are few tested, few labs..

Proxies that are closer..

If study ID is

Our systematic review and meta-analysis found a small positive increase in fitness benefits in the bird nests with increased green nest material, though the strength of these effects depended on the effect size metric used. SMDH showed a larger effect with the There is high heterogeneity among the effect sizes, emphasizing the context-dependence of this benefit. Importantly, our results reveal that the commonly studied mechanistic hypotheses of green nest material---whether through sexual signalling, parasite reduction, or direct nestling benefits---lack consistent empirical support across studies.

@@

------------------------------------------------------------------------

# Supplementary Appendix

@@

### PRISMA Checklist

Check with the online prisma list at <https://arc.net/l/quote/zhrgrlso>
